---
title: "CL_EDA_viz"
author: "Catherine Lucey"
date: "12/2/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(plotly)
library(patchwork)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

read in shsat data

```{r}

shsat_df = 
  read_csv("shsat_data.csv") %>% 
  select(-`...1`) %>% 
  mutate(
    year = factor(year, levels = c("2015", "2016", "2017", "2018", "2019", "2020"))
  )

```

load in extra ms directory data

```{r}

ms_dir_import =
  read_csv("raw_data/2019_DOE_Middle_School_Directory.csv") %>%
  janitor::clean_names() %>% 
  select(schooldbn, elaprof, mathprof) %>% 
  rename(dbn = schooldbn)

```


```{r}

boro_df =
  full_join(ms_dir_import, shsat_df, by = "dbn") %>% 
  drop_na() %>% 
  filter(n_offers > 5) %>% 
  mutate(
    test_offer_ratio = n_offers / n_testers
  )

n_offers_ggplot =
  boro_df %>% 
  group_by(year) %>% 
  ggplot(aes(x = year, y = n_offers)) +
  geom_boxplot() +
  facet_grid(. ~borough) +
  theme(axis.text.x = element_text(angle = 90))

n_testers_ggplot = 
  boro_df %>% 
  group_by(year) %>% 
  ggplot(aes(x = year, y = n_testers)) +
  geom_boxplot() +
  facet_grid(. ~borough) +
  theme(axis.text.x = element_text(angle = 90))

prop_offers_testers_ggplot =
  boro_df %>%
  mutate(
    offers_per_testers = n_offers / n_testers
  ) %>% 
  group_by(year) %>% 
  ggplot(aes(x = year, y = offers_per_testers)) +
  geom_boxplot() +
  facet_grid(. ~borough) +
  theme(axis.text.x = element_text(angle = 90))

# n_offers_ggplot / n_testers_ggplot / 
  
prop_offers_testers_ggplot / wt_boro_ggplot

```

NOTE: SHOULD PROBABLY SOMEHOW WEIGHT THE PROP TEST OFFERS VALUE BY THE NUMBER OF PEOPLE TAKING THE SHSAT

### below is the 6 year average (including 2020)

```{r}

avg_boro_df =
  boro_df %>% 
  group_by(dbn, borough) %>% 
  summarize(avg_n_offers = mean(n_offers),
            avg_n_testers = mean(n_testers),
            prop_test_offers = avg_n_offers / avg_n_testers)

bavg_n_offers_ggplot =
  avg_boro_df %>% 
  mutate(
    borough = as.factor(borough),
    borough = fct_reorder(borough, prop_test_offers)
  ) %>%
  ggplot(aes(x = borough, y = avg_n_offers)) +
  geom_boxplot()

bavg_n_testers_ggplot = 
  avg_boro_df %>% 
  mutate(
    borough = as.factor(borough),
    borough = fct_reorder(borough, prop_test_offers)
  ) %>%
  ggplot(aes(x = borough, y = avg_n_testers)) +
  geom_boxplot()

bavg_prop_offers_testers_ggplot =
  avg_boro_df %>% 
  mutate(
    borough = as.factor(borough),
    borough = fct_reorder(borough, prop_test_offers)
  ) %>%
  ggplot(aes(x = borough, y = prop_test_offers)) +
  geom_boxplot()

bavg_n_offers_ggplot / bavg_n_testers_ggplot / bavg_prop_offers_testers_ggplot + plot_layout(heights = c(2, 2, 2), widths = c(1, 1, 1))


```


## comparing 5 year pre-covid avg to the 2020 avg


```{r}

pre_covid_prop_offers_testers_ggplot =
  boro_df %>%
  filter(year != 2020) %>% 
  group_by(dbn, borough) %>% 
  summarize(avg_n_offers = mean(n_offers),
            avg_n_testers = mean(n_testers),
            prop_test_offers = avg_n_offers / avg_n_testers) %>% 
  mutate(
    borough = as.factor(borough),
    borough = fct_reorder(borough, prop_test_offers)
  ) %>%
  ggplot(aes(x = borough, y = prop_test_offers)) +
  geom_violin()

covid_prop_offers_testers_ggplot =
  boro_df %>%
  filter(year == 2020) %>% 
  group_by(dbn, borough) %>% 
  summarize(avg_n_offers = mean(n_offers),
            avg_n_testers = mean(n_testers),
            prop_test_offers = avg_n_offers / avg_n_testers) %>%
  mutate(
    borough = as.factor(borough),
    borough = fct_reorder(borough, prop_test_offers)
  ) %>%
  ggplot(aes(x = borough, y = prop_test_offers)) +
  geom_violin()

pre_covid_prop_offers_testers_ggplot + covid_prop_offers_testers_ggplot

```

what about the difference in proportions of offers/testers between precovid and covid?

```{r}
  
covid_prop_offers_testers =
  boro_df %>%
  filter(year == 2020) %>% 
  group_by(dbn, borough) %>% 
  summarize(avg_n_offers = mean(n_offers),
            avg_n_testers = mean(n_testers),
            covid_prop_test_offers = avg_n_offers / avg_n_testers) %>% 
  select(dbn, borough, covid_prop_test_offers)

pre_covid_prop_offers_testers = 
  boro_df %>%
  filter(year != 2020) %>% 
  group_by(dbn, borough) %>% 
  summarize(avg_n_offers = mean(n_offers),
            avg_n_testers = mean(n_testers),
            pre_prop_test_offers = avg_n_offers / avg_n_testers) %>% 
  select(dbn, borough, pre_prop_test_offers)

prop_diff_coivd_df = 
  full_join(covid_prop_offers_testers, pre_covid_prop_offers_testers, by = "dbn") %>%
  mutate(
    prop_diff = covid_prop_offers_testers - pre_covid_prop_offers_testers
  ) %>% 
  select(dbn, borough, prop_diff)

```

#### Looking at covid tests

```{r}

zip_mod = read_csv("raw_data/ZCTA-to-MODZCTA.csv") %>% 
  rename(zip = ZCTA)
covid_test <- read_csv("raw_data/tests-by-zcta.csv")

school_data = full_join(zip_mod, shsat_df, by = "zip") %>% 
  janitor::clean_names()

school_covidtest = full_join(school_data, covid_test, by = "modzcta") %>% 
  janitor::clean_names()


```

plotting

```{r}

covid_shsat_boro =
  school_covidtest %>%
  full_join(ms_dir_import, shsat_df, by = "dbn") %>%
  filter(n_offers > 5) %>% 
  filter(year == 2020) %>% 
  mutate(
    offers_per_testers = n_offers / n_testers
  ) %>% 
  select(dbn, borough, offers_per_testers, modzcta_cum_perc_pos)

covid_shsat_boro2 = 
  school_covidtest %>%
  full_join(ms_dir_import, shsat_df, by = "dbn") %>%
  filter(n_offers > 5) %>% 
  filter(year == 2020) %>%
  mutate(
    avg_n_testers = mean(n_testers),
    tester_wt = n_testers/avg_n_testers,
    test_offer_ratio = n_offers / n_testers,
    wt_offer_test_rt = test_offer_ratio*tester_wt
  )

two =
  covid_shsat_boro %>% 
  ggplot(aes(y = offers_per_testers, x = modzcta_cum_perc_pos, color = borough)) +
  geom_point()
one =
  covid_shsat_boro2 %>% 
    ggplot(aes(y = wt_offer_test_rt, x = modzcta_cum_perc_pos, color = borough)) +
    geom_point()

one / two


```



#### idk summary stats?

```{r}
shsat_zip_df %>%
  filter(n_offers > 5) %>%
  group_by(year) %>% 
  summarise(mean_offers = mean(n_offers),
            std_dev_offers = sd(n_offers))


shsat_zip_df %>% 
  filter(n_offers > 5) %>% 
   mutate(
    avg_n_offers = mean(n_offers),
    std_n_offers = n_offers / avg_n_offers
  ) %>% 
  filter(std_n_offers < 5) %>% 
  ggplot(aes(x = year, y = std_n_offers)) +
  geom_boxplot()

  select(year, n_offers) %>% 
  ggplot(aes(x = year, y = n_offers)) +
  geom_point()
  

    
```

separating by quartiles

```{r}

shsat_df %>% 
  mutate(
    offers_per_testers = n_offers / n_testers
  ) %>% 
  filter(n_offers > 5) %>% 
  summary()

shsat_df %>%
  filter(n_offers > 5) %>% 
  mutate(
    offers_per_testers = n_offers / n_testers
  ) %>%
  fabricatr::split_quantile(x = offers_per_tester, type = 4)
  

```

#### n testers vs n offers, accounting for ela and math proficiency

```{r}

boro_df %>% 
  group_by(year) %>% 
  ggplot(aes(x = n_testers, y = n_offers, color = borough)) +
  geom_point() +
  facet_grid(. ~year) +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") 

boro_df %>% 
  group_by(year) %>% 
  ggplot(aes(x = mathprof, y = test_offer_ratio)) +
  geom_point() +
  facet_grid(. ~borough) +
  theme(axis.text.x = element_text(angle = 90))

boro_df %>%
  filter(mathprof > 75) %>% 
  ggplot(aes(x = mathprof, y = n_offers, color = year)) +
  geom_point() +
  facet_grid(. ~borough)

# ela prof and math prof are correlated
boro_df %>%
  ggplot(aes(x = mathprof, y = elaprof, color = borough)) +
  geom_point() +
  facet_grid(. ~year)

```

### LOADING IN FULL DS and adding the ela data

```{r}

std_testing_df =
  read_csv("raw_data/2019_DOE_Middle_School_Directory.csv") %>%
  janitor::clean_names() %>% 
  select(elaprof, mathprof, schooldbn) %>% 
  rename(dbn = schooldbn) %>% 
  mutate(
    elaprof = as.numeric(elaprof),
    mathprof = as.numeric(mathprof)
  )

full_df = 
  read_csv("project_df.csv")
  
full_df =
  full_join(full_df, std_testing_df, by = "dbn") %>% 
  filter(n_offers > 5)
  

```

#### calculating weighted test prop ratio

```{r}

wt_boro =
  boro_df %>% 
  mutate(
    avg_n_testers = mean(n_testers),
    tester_wt = n_testers/avg_n_testers,
    wt_offer_test_rt = test_offer_ratio*tester_wt
  )

wt_boro_ggplot =
  wt_boro %>%
  group_by(year) %>% 
  ggplot(aes(x = year, y = wt_offer_test_rt)) +
  geom_boxplot() +
  facet_grid(. ~borough) +
  theme(axis.text.x = element_text(angle = 90))
  
    
  

```


```{r}

full_df %>% 
  weighted.mean(., n_offers, n_testers, na.rm = TRUE)

```


zip code averages

```{r}

wt_avgs_df =
  read_csv("filtered_weighted_avgs.csv") %>% 
  select(!`...1`) %>% 
  mutate(
    covid_diff_testers = `2015-19_avg_testers` - avg_testers_2020,
    covid_diff_offers = `2015-19_avg_offers` - avg_offers_2020 ,
    covid_diff_wt_avg = `2015-19_wt_avg` - wt_avg_2020
  )

pre_gg =
  wt_avgs_df %>% 
  ggplot(aes(x = `2015-19_wt_avg`)) +
  geom_histogram()

`2020_gg` = 
  wt_avgs_df %>% 
  ggplot(aes(x = `wt_avg_2020`)) +
  geom_histogram()

wt_avg_diff_gg =
  wt_avgs_df %>% 
  ggplot(aes(y = covid_diff_wt_avg, x = modzcta)) +
  geom_point()

tester_avg_diff_gg =
  wt_avgs_df %>% 
  ggplot(aes(y = covid_diff_testers, x = modzcta)) +
  geom_point()

offer_avg_diff_gg = 
  wt_avgs_df %>% 
  ggplot(aes(y = covid_diff_offers, x = modzcta)) +
  geom_point()

pre_gg / `2020_gg`

wt_avg_diff_gg + offer_avg_diff_gg + tester_avg_diff_gg

```

chi sq to see if the delta per school is significantly different between 2015-2019 and 2020

```{r}

chisq_df =
  read_csv("project_df.csv") %>%
  filter(n_offers > 5) %>% 
  select(modzcta, dbn, n_testers, n_offers, year) %>%
  drop_na() %>% 
  group_by(dbn, year) %>% 
  pivot_wider(names_from = year, values_from = n_testers:n_offers) %>%
  drop_na() %>% 
  summarize(
    `2015-19_avg_testers` = mean(n_testers_2015:n_testers_2019),
    `2020_testers` = n_testers_2020,
    `2015-19_avg_offers` = mean(n_offers_2015:n_offers_2019),
    `2020_offers` = n_offers_2020
  ) %>% 
  mutate(
    `2015-19_prop` = `2015-19_avg_offers` / `2015-19_avg_testers`,
    `2020_prop` = `2020_offers` / `2020_testers`
  )

chisq_df %>% 
  select(`2015-19_avg_testers`, `2020_testers`) %>% 
  chisq.test()

chisq_df %>% 
  select(`2015-19_avg_offers`, `2020_offers`) %>% 
  chisq.test()

chisq_df %>% 
  select(`2015-19_prop`, `2020_prop`) %>%
  chisq.test()

%>%
  prop.test(`2015-19_prop`, `2020_prop`, p = NULL, alternative = "two.sided")

 mutate(
    `2015-19_prop` = `2015-19_prop` * 100,
    `2020_prop` = `2020_prop` * 100
  ) %>% 

```

plotting stuff from that chisq dataset

