---
title: "SHSAT_clean_CML"
author: "Catherine Lucey"
date: "11/23/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

### Description

This RMD includes code to import, clean, and join the datasets of SH offers by middle school and middle school locations + descriptors.

### Variable dictionary

dbn = dbn (district borough number) of the sending middle school (original = feeder_school_dbn)

feeder_school = name of the middle school sending students (original = feeder_school_name)

n_offers = number of specialized high school offers per middle school (original = count_of_offers (2015-2017) or number_of_offers (2018 onward))

n_testers = number of students taking the SHSAT per middle school (original = count_of_testers)

n_hs_applicants = number of students per middle school applying out to high school (not just specialized high schools) (original = count_of_students_in_hs_admissions)

ms_info_df: ms = "middle school"

#### Notes

SHSAT data were downloaded on 11/23/21 at 3:30PM. API links are:
https://data.cityofnewyork.us/resource/xqx4-kdvp.csv,
https://data.cityofnewyork.us/resource/8ws3-956v.csv,
https://data.cityofnewyork.us/resource/vsgi-eeb5.csv,
https://data.cityofnewyork.us/resource/uf53-ree9.csv,
https://data.cityofnewyork.us/resource/xuij-x4t4.csv,
https://data.cityofnewyork.us/resource/k8ah-28f4.csv

```{r setup, include=FALSE}

library(tidyverse)
library(rvest)
library(httr)
library(readxl)

```

```{r, load shsat data}

shsat_2_df =
  tibble(
    file_name = list.files("raw_data/shsat_2")
  ) %>% 
  mutate(
    path = str_c("raw_data/shsat_2/", file_name),
    data = map(path, read_csv),
    data = map(data, janitor::clean_names),
    data = map(data, as_tibble),
    year = str_remove(file_name, "_SHSAT_Admissions_Test_Offers_By_Sending_School.csv"),
  ) %>% 
  select(year, data) %>% 
  bind_rows() %>% 
  unnest(cols = c(data))

%>% 
  unnest(data)

shsat_1_df[[2]][[3]]

tidy_shsat = function(column_name) {
  if ("count" %in% column_name) return(as.double(column_name))
}

  mutate
  return(column_name)

if (str_detect(column_name, "count") == TRUE) return(as.double(column_name))

tidied = tidy_shsat(column_name = shsat_1_df$data)

str_detect(shsat_1_df$data)


#shsat_df_combined$data[[1]]

shsat_df =
  GET("https://data.cityofnewyork.us/resource/k8ah-28f4.csv") %>% 
  content("parsed") 

```

### Idea 1: Importing and Cleaning SHSAT Data from 2015-2016 to 2017-2018

```{r}

file_names_df =
  tibble(
    file_name = list.files("raw_data/shsat_1")
  )

load_shsat = function(file_name) {
  
  path = str_c("raw_data/shsat_1/", file_name)
  
  data = read_csv(path)
  
  year = str_remove(file_name, "_SHSAT_Admissions_Test_Offers_By_Sending_School.csv")
  
  clean_df = 
    data %>% 
    janitor::clean_names() %>% 
    cbind(year) %>% 
    rename(
      dbn = feeder_school_dbn,
      feeder_school = feeder_school_name,
      n_offers = count_of_offers,
      n_testers = count_of_testers,
      n_hs_applicants = count_of_students_in_hs_admissions,
      ) %>% 
    mutate(
      n_hs_applicants = 
        replace(n_hs_applicants, n_hs_applicants == "0-5", "5"),
      n_hs_applicants = as.double(n_hs_applicants)
    ) %>% 
    mutate(
      n_offers = 
        replace(n_offers, n_offers == "0-5", "5"),
      n_offers = as.double(n_offers)
    ) %>% 
    mutate(
      n_testers =
        replace(n_testers, n_testers == "0-5", "5"),
      n_testers = as.double(n_testers)
    )
  
  return(clean_df)
}

```

#### Generating Two Functions for Importing and Cleaning SHSAT Data from 2015-2016 to 2019-2021

```{r}

clean_shsatv1 = function(df){
  clean_df = 
    df %>% 
    janitor::clean_names() %>% 
    rename(
      dbn = feeder_school_dbn,
      feeder_school = feeder_school_name,
      n_offers = count_of_offers,
      n_testers = count_of_testers,
      n_hs_applicants = count_of_students_in_hs_admissions,
      ) %>% 
    mutate(
      n_hs_applicants = 
        replace(n_hs_applicants, n_hs_applicants == "0-5", "5"),
      n_hs_applicants = as.double(n_hs_applicants)
    ) %>% 
    mutate(
      n_offers = 
        replace(n_offers, n_offers == "0-5", "5"),
      n_offers = as.double(n_offers)
    ) %>% 
    mutate(
      n_testers =
        replace(n_testers, n_testers == "0-5", "5"),
      n_testers = as.double(n_testers)
    )
  
  return(clean_df)
}

clean_shsatv2 = function(df){
  clean_df =
    df %>% 
    janitor::clean_names() %>% 
    rename(
      dbn = feeder_school_dbn,
      feeder_school = feeder_school_name,
      n_offers = number_of_offers,
      n_testers = count_of_testers,
      n_hs_applicants = count_of_students_in_hs_admissions,
      ) %>% 
    mutate(
      n_hs_applicants = 
        replace(n_hs_applicants, n_hs_applicants == "0-5", "5"),
      n_hs_applicants = as.double(n_hs_applicants)
    ) %>% 
    mutate(
      n_offers = 
        replace(n_offers, n_offers == "0-5", "5"),
      n_offers = as.double(n_offers)
    ) %>% 
    mutate(
      n_testers =
        replace(n_testers, n_testers == "0-5", "5"),
      n_testers = as.double(n_testers)
    )
  
  return(clean_df)
}

```

#### Loading Data Across Years

```{r}

fifteen_df = 
  read_csv("raw_data/shsat_1/2015-2016_SHSAT_Admissions_Test_Offers_By_Sending_School.csv") %>% 
  mutate(year = "2015") %>% 
  clean_shsatv1()

sixteen_df =
  read_csv("raw_data/shsat_1/2016-2017_SHSAT_Admissions_Test_Offers_By_Sending_School.csv") %>% 
  mutate(year = "2016") %>% 
  clean_shsatv1()

seventeen_df =
  read_csv("raw_data/shsat_1/2017-2018_SHSAT_Admissions_Test_Offers_By_Sending_School.csv") %>% 
  mutate(year = "2017") %>% 
  clean_shsatv1()

eighteen_df =
  read_csv("raw_data/shsat_1/2018-2019_SHSAT_Admissions_Test_Offers_By_Sending_School.csv") %>% 
  mutate(year = "2018") %>% 
  clean_shsatv2()

nineteen_df =
  eighteen_df =
  read_csv("raw_data/shsat_1/2019-2020_SHSAT_Admissions_Test_Offers_By_Sending_School.csv") %>% 
  mutate(year = "2019") %>% 
  clean_shsatv2()

twenty_df =
  read_csv("raw_data/shsat_1/2020-2021_SHSAT_Admissions_Test_Offers_By_Sending_School.csv") %>% 
  mutate(year = "2020") %>% 
  clean_shsatv2()

```

#### Joining and Cleaning SHSAT Datasets: Removing Missing/Inconsistent Values

```{r}

shsat_df = 
  rbind(
    fifteen_df, 
    sixteen_df, 
    seventeen_df, 
    eighteen_df, 
    nineteen_df, 
    twenty_df) %>%
  unique()

```


#### Importing and Cleaning the MS Directory Datafile

```{r}

ms_dir_import =
  read_csv("raw_data/2020_DOE_Middle_School_Directory.csv") %>% 
  janitor::clean_names() %>% 
  select(schooldbn, borough, postcode, address) %>% 
  rename(
    dbn = schooldbn,
    zip = postcode
  )


ms_dir_df =
  separate(address, into = c("address", "zip"), sep = "NY")
  
ms_zip_dbn =
  ms_dir_df %>% 
  select(dbn, zip, borough)

```

substr(start = -5

https://stackoverflow.com/questions/7963898/extracting-the-last-n-characters-from-a-string-in-r?rq=1
https://stringr.tidyverse.org/reference/str_sub.html

< 100 schools is too low

unique fxn --> specify w/in that x = dbn to just get the unique middle schools 

jk 

pivot wider --> drop nas, then pivot wider again to make the year variable

https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/unique

convert the years to one year



#### Joining Zipcode data to SHSAT

3620 observations before dropping na values, 2789 obs after dropping na values --> should be ~465 schools

```{r}

shsat_zip_df =
  full_join(shsat_df, ms_dir_import, by = "dbn") %>% 
  drop_na() %>% 
  mutate(
    year = factor(year, levels = c("2015", "2016", "2017", "2018", "2019", "2020"))
  )

shsat_zip_df %>% 
  group_by(year) %>% 
  select(zip) %>% 
  count(is.na(.))

  filter(borough == "N/A") %>% 
  view()
  count()
  

```

#### Export SHSAT Data

```{r}

write.csv(shsat_zip_df, "shsat_data.csv")

```


#### figuring out shared building nonsense

```{r}

shared_df =
  ms_dir_df %>% 
  filter(sharedbuilding != "No")

```


#### DATA!!!! (eda)

will have to separate into quartiles and look at variation within quartiles

```{r}

shsat_zip_df %>%
  filter(n_offers > 5) %>%
  group_by(year) %>% 
  summarise(mean_offers = mean(n_offers),
            std_dev_offers = sd(n_offers))

shsat_zip_df %>% 
  filter(n_offers > 5) %>% 
  summary()


shsat_zip_df %>% 
  filter(n_offers > 5) %>% 
   mutate(
    avg_n_offers = mean(n_offers),
    std_n_offers = n_offers / avg_n_offers
  ) %>% 
  filter(std_n_offers < 5) %>% 
  ggplot(aes(x = year, y = std_n_offers)) +
  geom_boxplot()

  select(year, n_offers) %>% 
  ggplot(aes(x = year, y = n_offers)) +
  geom_point()

```

